Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population

### 2022-09-25 src/tex/appendicies.tex
methods.tex is mostly finished but w/o citations
cleanup more of appendices/methods

We explicitly enumerate each of the models used in this work for $p(\theta|\Lambda)$, along with 
their respective hyper-parameters and prior distributions in the next section. To calculate marginal likelihoods and draw 
samples of the hyper parameters from the hierarchical posterior distribution shown in equation \ref{importance-posterior}, we use the 
\textsc{Bilby} \citep{Ashton_2019, bilby_gwtc1} and \textsc{GWPopulation} \citep{Talbot_2019} Bayesian inference software libraries with 
the \textsc{Dynesty} dynamic nested sampling algorithm \citep{Speagle_2020}. We also make use of the hamiltonian monte carlo algorithms in 
\textsc{numpyro} and use \textsc{jax} to calculate our likelihoods \bruce{FIND citations for these}.

\section{Models and Hyper-Parameters} \label{sec:hypparams}

>>>>

We explicitly enumerate each of the models used in this work for $p(\theta|\Lambda)$, along with 
their respective hyper-parameters and prior distributions in the next section. To calculate draw 
samples of the hyper parameters from the hierarchical posterior distribution shown in equation \ref{importance-posterior}, we use the 
NUTS hamiltonian monte carlo sampler in \textsc{numpyro} and \textsc{jax} to calculate likelihoods \bruce{FIND citations for these}.

\section{Penalized Splines and Smoothing}\label{sec:psplines}

Spline functions have been shown to be sensitivite to the chosen number of knots, and their locations or spacing. 
Adding more knots knots increases the a-priori variance in the spline function, while the space between knots can limit the 
resolution of features in the data the spline is capable of resolving. To ensure your spline based model is flexible enough 
one would want to add as many knots as densely as possible, but this comes with the inherit larger variance as mentioned above. 
To overcome this we resort to the frequentist penalized spline (P-Spline) in which one applies a penalty or regularization term 
to the likelihood based on the difference of adjacent knot coeficients. The linear combination of spline basis components is 
is flat when the basis coeficients are equal (see Figure \ref{fig:spline_basis}). By penalizing the likelihood as the differences between adjacent
knot coeficients get larger, one gets a smoothing of the spline function inferences. Since we are using hierarchical Bayesian inferece as our 
statistical framework, we formulate this penalized likelihood with its Bayesian analog. The difference penalty is equivalent to applying a stochastic 
differnece prior among the knots, as described in \bruce{CITE THIS PAPER}. This is also sometimes referred to as a Gaussian random walk prior, since
each coeficient is normally distributed about its neighboring coeficients. We then use as our smoothing prior, for a spline basis with $n$ degree's of freedom
, and a difference order of $r$ in our smoothing penalty (see \bruce{cite P-splines paper}) is defined as:

\begin{eqnarray}
\tau_\lambda \sim \Gamma(a, b), \\
p(\bm{c} | \tau_\lambda) \propto \exp \big[ -\frac{1}{2} \tau_\lambda \bm{c}^{\mathrm{T}} \bm{D_r}^{\mathrm{T}} \bm{D_r} \bm{c}  \big] e^{-n\bm{c}}
\end{eqnarray}

\noindent Above $\bm{D_r}$ is the order-$r$ difference matrix, of shape $(n-r \times n)$ and $a$, $b$, shape parameters for the Gamma distribution. 
This smoothing prior removes the strong dependence on number and location of knots that arises with using splines. By adding in a very large number 
of knots such that your domain is densly populated with basis coeficients, this allows the freedom for the model to find the smoothing 
scale that the data prefers. One downside of this prior specification as noted by\ bruce{CITE BAYESIAN PSPLINES}, is a slight dependence on choice 
of the b hyper-parameter for Gamma distribution prior on $\tau_\lambda$. We reduce this affect by following the method introduced in \bruce{CITE BAYESIAN PSPLINES} 
to alleviate this. This is done by modeling $\tau_\lambda$ with a mixture model on gamma distributions with varying b hyper-parameters. 
We now replace the prior on $\tau_\lambda$ with:

\begin{align}
    \bm{p} \sim \mathcal{D}(m), \\
    p(\tau_\lambda | \bm{p}) = \sum_i^m p_i \mathcal{G}(1, b_m)
\end{align}

%Additionally one can implement a spatially adaptive smoothing prior which allows for a smoothing prior
%that can smoothly evolve over the domain of your spline function. This is done in practice by having the $\tau_\lambda$ parameter evolve smoothly
%across the knots. \bruce{CITE BAYESIAN PSPLINES}, formulates a spatially adaptive prior in this way by adding in $n-r-1$ additional parameters that
%effectively modulate the $\tau_\lambda$ across each difference. We now replace our prior on the coeficients with:

%\begin{eqnarray}
%\bm{\lambda} \sim \mathcal{G}(\omega, \omega), \\
%\bm{\Lambda} \equiv \mathrm{diag}\big(1, \lambda_0,...,\lambda_{n-r-1}\big), \\
%p(\bm{c} | \tau_\lambda) \propto \exp \big[ -\frac{1}{2} \tau_\lambda \bm{c}^{\mathrm{T}} \bm{D_r}^{\mathrm{T}} \bm{\Lambda} \bm{D_r} \bm{c} \big] e^{-n\bm{c}}
%\end{eqnarray}

%Where above the matrix $\bm{\Lambda}$ a diagonal matrix of shape $(n-r \times n-r)$, and $\omega$ an arbitrarily small value such that the Gamma distribution 
%prior on $\bm{\lambda}$ has a mean of 1 and an arbitrarily large variance. This enforces a smooth evolution of the smoothing penalty 
%across the domain that is being analyzed. This type of prior is especiially important when modeling distributions that span large orders of magnitude like 
%the power-law like mass distributions that are commonly used.

Since the smoothing prior is really wanting to pull the coeficients closer to each other to smooth the distribution, 
we must ensure that the spline function is in fact flat or smooth with all equal coeficients. Often times one would stack knots at the same values 
at hard parameter boundaries but this removes the feature that with equal weights the spline function is flat. We follow the distinction from \bruce{CITE psplines} 
that the difference penalty or prior is only valid with ``proper'' spline bases. A proper basis is where all of the knots are evenly spaced, see Figure \ref{fig:spline_basis}.

\section{Model specification} \label{sec:modelpriors}
